---
title: "LotzLab2"
author: "Justin Lotz"
date: "18/05/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Working with vectors

### Question 1a
```{r}
x1 <- seq(2, 10, 2)

```

### Question 1b
```{r}
x2 <- x1[-4]

```

### Question 1c
```{r}
x3 <- seq(107.5, 109.5, 0.5)

```

### Question 1d
```{r}
x4 <- rep(seq(11, 33, 11), times = 2)

```

### Question 1e
```{r}
x5 <- rep(c(11, 22, 33), each = 2)

```

### Question 1f
```{r}
x6 <- c(rep(3, 2), rep(1, 3), -1)

```

### Question 1g
```{r}
x7 <- cumsum(4:7)

```

### Question 1h
```{r}
x8 <- cumprod(seq(-3, 5, 2))

```

## Question 2: Working with Indexes

### Question 2a
```{r}
x10 <- x1[1:2]

```

### Question 2b
```{r}
x11 <- x3[c(1, 4)]

```

### Question 2c
```{r}
x12 <- x3[c(4, 1)]

```

### Question 2d
```{r}
x13 <- x3[seq(5, 1, -1)]

```

### Question 2e
```{r}
x14 <- x7[c(3, 4, 1, 2, 1)]

```

## Analyzing SACU GDP data

### Question 3: Load File into R Studio code chunk
```{r}
load(file="sacuGDP.RData")

```

### Question 4: Use an index to find the most recent GDP figure for South Africa
```{r}
# GPD Per country assignment to dataframe
df <- data.frame(bwa, lso, nam, swz, zaf)

# Obrtain latest ZAF GDP figures
df$zaf[15]

```

### Question 5: Use a function to find to find the lowest GDP figure for Swaziland
```{r}
min(df$swz)

```

### Question 6: Use a function (or two) to figure out whether Lesotho’s GDP ever declined year-on-year during the period 2000–2014.
```{r}
# Use diff function to identify any negative differences. There are none, therefore no decline took place.
diff(df$lso)

```

### Question 7:  Use R to figure out which country had the second-highest average GDP over the decade
```{r}
# Assign mean to gdp.mean
gdpMeans <- c(mean(bwa), mean(lso), mean(nam), mean(zaf), mean(swz))

names(gdpMeans) <- c("bwa", "lso", "nam", "swz", "zaf")

```

### Question 8: Use vector arithmetic to calculate a vector called sacu that contains the total GDP for SACU in each year.
```{r}
sacu <- bwa + lso + nam + swz + zaf

```

### Question 9: Enter the following command: names(sacu) <- 2000:2014
```{r}
names(sacu) <- 2000:2014

```

### Question 10: Make a “line” plot of sacu
```{r}
plot(sacu, type = "l", col = "red")

```

### Question 11: If your plot command was successful, enter the following three commands in a code chunk
```{r}
pdf(file="lab2_SACUplot.pdf")
plot(sacu, type = "l", col = "red")
dev.off()

```



